---
title: "Acute Stress Reduction Analysis: Music vs. Podcast Intervention"
author: "Frederico Pedrosa"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
---

Data analysis for an experiment that investigated whether listening to music (experimental group) for 25 minutes versus listening to a podcast (control group) for 25 minutes reduces participants' stress levels

## **Preparing data**

```{r setting}
library(readxl)     #read .xlsx files
library(haven)      #treat data
library(ggplot2)    #draw images
library(gridExtra)  #treat images
library(nparLD)     #non parametric ANOVA
library(dplyr)      #allow code in pipelines
library(rstatix)    #estimate effect sizes

dado <- read_excel("~/Isabela/Dados_Agudo.xlsx")
dado$Grupo <- ifelse(dado$Grupo == "Controle", "Control", "Experimental")
dado <- as_factor(dado)
table(dado$Grupo)
```

**Preparing salivary cortisol data**

The values for salivary cortisol (Cortpre and Cortpos) were multiplied by 27.7 to convert the units from nmol/L to ng/dL. This conversion is necessary because some reference ranges and comparative studies report cortisol levels in ng/dL. The factor 27.7 is derived from the molecular weight of cortisol (approximately 362.5 g/mol), allowing for consistent interpretation and comparison of results across different measurement systems.

The data do not follow a normal distribution.

```{r}
dado$Cortpre <- dado$Cortpre * 27.7
dado$Cortpos <- dado$Cortpos * 27.7
shapiro.test(dado$Cortpre)
shapiro.test(dado$Cortpos)

```


```{r}
# Subset data by group
intervention_data <- dado[dado$Grupo == "Experimental", ]
control_data <- dado[dado$Grupo == "Control", ]

# Create a combined boxplot and scatter plot for Cortpre in the Experimental group
combined_plot_intervention_pre <- ggplot(intervention_data, aes(x = Grupo, y = Cortpre)) +
  geom_boxplot(position = position_dodge(width = 0.8), color = "black", fill = "white") +
  geom_jitter(position = position_jitter(width = 0.2), size = 2) +
  labs(title = "Before",
       x = "",
       y = "Cortisol Level (µg/dL)") +
  ylim(0, 45) +
  scale_x_discrete(labels = c("Experimental")) +
  theme(legend.position = "none")

# Create a combined boxplot and scatter plot for Cortpos in the Experimental group
combined_plot_intervention_pos <- ggplot(intervention_data, aes(x = Grupo, y = Cortpos)) +
  geom_boxplot(position = position_dodge(width = 0.8), color = "black", fill = "white") +
  geom_jitter(position = position_jitter(width = 0.2), size = 2) +
  labs(title = "After",
       x = "",
       y = "Cortisol Level (µg/dL)") +
  ylim(0, 45) +
  scale_x_discrete(labels = c("Experimental")) +
  theme(legend.position = "none")

# Create a combined boxplot and scatter plot for Cortpre in the Control group
combined_plot_control_pre <- ggplot(control_data, aes(x = Grupo, y = Cortpre)) +
  geom_boxplot(position = position_dodge(width = 0.8), color = "black", fill = "white") +
  geom_jitter(position = position_jitter(width = 0.2), size = 2) +
  labs(title = "Before",
       x = "",
       y = "Cortisol Level (µg/dL)") +
  ylim(0, 45) +
  scale_x_discrete(labels = c("Control")) +
  theme(legend.position = "none")

# Create a combined boxplot and scatter plot for Cortpos in the Control group
combined_plot_control_pos <- ggplot(control_data, aes(x = Grupo, y = Cortpos)) +
  geom_boxplot(position = position_dodge(width = 0.8), color = "black", fill = "white") +
  geom_jitter(position = position_jitter(width = 0.2), size = 2) +
  labs(title = "After",
       x = "",
       y = "Cortisol Level (µg/dL)") +
  ylim(0, 45) +
  scale_x_discrete(labels = c("Control")) +
  theme(legend.position = "none")

# Display the plots side by side
(grid_arrangement <- grid.arrange(combined_plot_intervention_pre, combined_plot_intervention_pos, combined_plot_control_pre, combined_plot_control_pos, ncol = 2))
```


## **Transforming data em long style**


```{r }
dado_longo <- read_excel("~/Isabela/dado_longo.xlsx")
dado_longo$Cort <- dado_longo$Cort * 27.7



dado_longo <- dado_longo %>%
  mutate(
    Grupo = recode(Grupo,
                   "Controle" = "Control",
                   "Intervenção" = "Experimental"),
    Tempo = recode(Tempo,
                   "1" = "Before",
                   "2" = "After")
  ) %>%
  rename(
    Group = Grupo,
    Time = Tempo
 )


dado_longo <- dado_longo %>%
  mutate(ID = rep(1:(nrow(dado_longo) / 2), each = 2))

dado_longo$Grupo <- as.factor(dado_longo$Group)
dado_longo$Tempo <- as.factor(dado_longo$Time)

```

## **Nonparametric Factorial Repeated Measures ANOVA**

There are no diferences beetween the to groups.

```{r }
mod_1x1 <- nparLD(formula = Cort ~ Time*Group, # Fórmula como ANOVA 
                  data = dado_longo, 
                  subject= dado_longo$ID, # Variável do sujeito
                  description = F  # Descrição da análise
                  
)

summary(mod_1x1)
```

There is no difference beetween groups over time, but there are within-group effects.
Besides that, we can investigate the magnitude of this difference internally.

```{r}
data_wilcox_exp <- data.frame(
  Cortisol_pre = intervention_data$Cortpre,
  Cortisol_pos = intervention_data$Cortpos
)

data_wilcox_exp <- stack(data_wilcox_exp)

# Executar o teste de Wilcoxon
wilcox_result_exp <- wilcox.test(values ~ ind, data = data_wilcox_exp)

# Imprimir os resultados
print(wilcox_result_exp)

effect <- wilcox_effsize(values ~ ind, data= data_wilcox_exp, paired = TRUE, 
                         ci=TRUE, ci.type= "bca", nboot= 1000)
effect
```


```{r}
# Criar um objeto de dados
data_wilcox_C <- data.frame(
  Cortisol_pre = control_data$Cortpre,
  Cortisol_pos = control_data$Cortpos
)

# Criar um objeto de dados no formato longo a partir de dados_wilcox
data_wilcox_C <- stack(data_wilcox_C)

# Executar o teste de Wilcoxon
wilcox_result_C <- wilcox.test(values ~ ind, data = data_wilcox_C)

# Imprimir os resultados
print(wilcox_result_C)

effectC <- wilcox_effsize(values ~ ind, data= data_wilcox_C, 
                          ci=TRUE, ci.type= "bca", nboot= 1000)
effectC

```

The effect is larger in the experimental group.

```{r}
sessionInfo()
```


